Leadership across cultures and contexts: an exploratory analysis

### 2020-04-26 leadership_across_cultures_contexts.Rmd
Looking at function and quality variable support by subsistence and region.

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by mode of subsistence and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture. Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis. Symbol shapes indicate mode of subsistence and symbol size indicates the number of documents available for that culture."}
library(maps)
world_map <- map_data("world")
p_map <- ggplot(world_map, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill="lightgray", colour = "white") +
  geom_point(
    data = leader_cult, 
    aes(
      c_longitude, 
      c_latitude, 
      colour = subsistence, 
      shape = subsistence, 
      size = documents
    )
  ) +
  coord_fixed() +
  labs(x='', y='', shape = 'Subsistence', colour = 'Subsistence', size = 'Documents') +
  guides(shape = guide_legend(override.aes = list(size = 4))) +
  theme_bw(15)
ggsave(filename = 'Figures/map.pdf', p_map)
p_map
```

@@

## Distribution of text records by mode of subsistence and group type

The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups (see Methods section). Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by mode of subsistence. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

@@

```{r subsisttree, fig.cap="The area of the entire rectangle represents the total number of text records in the study data (N = 2012). The nested rectangles represent (from smallest to largest) the proportion of text records in each document, culture, and mode of subsistence."}

@@

## Continental region is confounded with mode of subsistence

@@

```{r highstatusplot, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record. Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010]. Blue values are positive predictors; red values are negative predictors. X-axis is on a log 10 scale."}
highstatus_plot + scale_x_log10()
```

@@

@@

We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`. We then fit separate logistic mixed effects models of each feature as a function of subsistence mode, with random effects for author nested within culture. After adjusting the subsistence coefficient p-values for an FDR of 0.05, we computed and plotted the estimated marginal means of the remaining significant model(s). See Figure \@ref(fig:features).

```{r features, fig.width=8, fig.cap="Esimtated marginal means [@lenth2020] of logistic mixed effects regression models of Feature variables by mode of subsistence. Only models whose subsistence coefficient had a corrected p-value < 0.10. Plotted on the response scale. See text for details."}

# x <- pmap(sub_models_sig, ~ plot(..6, main = ..1))

plot_feature_models

>>>>

The geographic distribution of the cultures in our sample is displayed in Figure \@ref(fig:geographic), with cultures labeled by subsistence strategy and represented by a symbol proportional to the number of documents that provided ethnographic texts for that culture. Our search yielded `r nrow(all_data)` text records from `r length(unique(leadershipdata::documents$d_ID))` unique documents describing `r length(unique(all_data$d_culture))` cultures. The number of text records per culture ranged from `r min(leader_cult$number_leader_records, na.rm=T)` to `r max(leader_cult$number_leader_records, na.rm=T)`, with a median of `r median(leader_cult$number_leader_records, na.rm=T)`, a mean of `r signif(mean(leader_cult$number_leader_records, na.rm=T), 3)`, and a standard deviation of `r signif(sd(leader_cult$number_leader_records, na.rm=T), 3)`. The text records were generally short, ranging from `r textstats$min` to `r textstats$max` words, with a median word count of `r textstats$median`, a mean of `r textstats$mean`, and standard deviation of `r textstats$sd`. Of the `r nrow(all_data)` text records, `r final_record_count` had evidence for at least one variable in the study.

@@

```{r geographic, fig.width=15, fig.cap="The geographical distributions of cultures in this analysis. Symbol shapes indicate subsistence strategy and symbol size indicates the number of documents available for that culture."}
library(maps)
world_map <- map_data("world")
p_map <- ggplot(world_map, aes(x = long, y = lat)) +
  geom_polygon(aes(group = group), fill="lightgray", colour = "white") +
  geom_point(
    data = leader_cult, 
    aes(
      c_longitude, 
      c_latitude, 
      colour = subsistence, 
      shape = subsistence, 
      size = documents
    )
  ) +
  coord_fixed() +
  labs(x='', y='', shape = 'Subsistence', colour = 'Subsistence', size = 'Documents') +
  guides(shape = guide_legend(override.aes = list(size = 4))) +
  theme_bw(15)
ggsave(filename = 'Figures/map.pdf', p_map)
p_map
```

@@

## Distribution of text records by subsistence strategy and group type

The type of group to which a leader text referred was coded into seven researcher-coded categories: residential groups, kin groups, economic groups, political groups at the community level, political group beyond the community level, military groups, and religious groups (see Methods section). Figure \@ref(fig:grouptype) indicates the proportion of text records that were categorized into each of these groups, broken out by subsistence strategy. We rely on a modified classification of subsistence from the HRAF, collapsing “hunter-gatherers” and “primarily hunter-gatherers” into a single category, “hunter-gatherers,” and by collapsing “agro-pastoralists” and “pastoralists” into a single “pastoralists” category. We renamed the HRAF’s “Other” subsistence category, which is applied to societies with various combinations of hunting, fishing, gathering, pastoralism, and horticulture or intensive agriculture, as “mixed.” This produced a subsistence classification with five groups: Hunter-gatherers, Pastoralists, Horticulturalists, Agriculturalists, and Mixed.

@@

```{r subsisttree, fig.cap="The area of the entire rectangle represents the total number of text records in the study data (N = 2012). The nested rectangles represent (from smallest to largest) the proportion of text records in each document, culture, and subsistence strategy."}

@@

## Continental region is confounded with subsistence strategy

@@

```{r highstatusplot, echo=F, message=F, fig.width=8, fig.height=6, fig.align='center', fig.cap="Lasso logistic regression coefficients, which indicate the words whose frequencies in each text record best predicted evidence for the *high status* leader quality in each text record. Coefficients estimated with glmnet package using cross-validation to determine the optimal value of lambda [@friedman_regularization_2010]. Red values are positive predictors; purple values are negative predictors. X-axis is on a log 10 scale."}
highstatus_plot + scale_x_log10()
```

@@

## Effects of subsistence strategy and region on function and quality variables

```{r varsubsisregion, fig.cap="Estimated marginal means of logistic mixed effects models of each leader function and quality variable as a function of subsistence strategy and region, with random effects for author nested within culture. These are the only models with significant coefficients for either subsistence strategy or region, after correcting p-values for an FDR of 0.05."}

plot_morality_region / plot_prosocial_region
```

@@

We constructed `r length(features)` "feature" variables from the higher-level clusters (Figure \@ref(fig:cluster)): `r paste(features, collapse = ', ')`. We then fit separate logistic mixed effects models of each feature as a function of subsistence strategy and region, with random effects for author nested within culture. After adjusting the subsistence coefficient p-values for an FDR of 0.05, we computed and plotted the estimated marginal means of the significant models. See Figure \@ref(fig:features).

```{r features, fig.width=11, fig.cap="Esimtated marginal means [@lenth2020] of logistic mixed effects regression models of Feature variables by subsistence strategy and region. Only models whose subsistence coefficient had a corrected p-value < 0.05. The coefficient for region was also significant in the model of the 'Big man' feature. Plotted on the response (probablity) scale. See text for details."}

feature_scale <- scale_x_continuous(limits = c(0, 0.20))

p_subsis_mediate <- ggemmeans(subsis_models_sig$emmeans_subsis[[1]]) +
  feature_scale +
  labs(title = 'Mediate', x = '', y = '')

p_subsis_bigman <- ggemmeans(subsis_models_sig$emmeans_subsis[[2]]) +
  feature_scale +
  labs(title = 'Big man', x = '', y = '')

p_region_mediate <- ggemmeans(subsis_models_sig$emmeans_region[[1]]) +
  feature_scale +
  labs(title = 'Mediate', x = '', y = '')

p_region_bigman <- ggemmeans(subsis_models_sig$emmeans_region[[2]]) +
  feature_scale +
  labs(title = 'Big man', x = '', y = '')

(p_subsis_mediate + p_region_mediate) / (p_subsis_bigman + p_region_bigman)
